Communicating human related spatial distributions: A comparison of spatial visualisations using visual inference.

### 2020-01-07 paper/paper.Rmd

```{r hist_height, fig.cap = "Six histograms show the distribution of the time taken to submit a response for each combination of trend and type of display. The time taken to evaluate each display is broken into five second windows. The height of the histogram bars show how many evaulations were submitted within each time window. The orange regions of the bars show the amount of correct detections, and the green regions are the amount of incorrect detections."}
ggplot(d %>% rename(Detected = detect_f), 
  aes(x = time_taken, fill = Detected, group = Detected)) + 
  scale_fill_manual(values = detect_colours) +
  geom_histogram(binwidth = 5, boundary = 0) +
  facet_grid(type ~ trend) + 
  theme(legend.position = "bottom") +
  labs(x = "Time taken (Seconds)", y = "Amount of evaluations")
```

```{r hist_fill, fig.cap = "The time taken to evaluate each display is broken into five second windows. The height of the histogram bars show the proportion of evaulations that were submitted within each time window."}
ggplot(d %>% rename(Detected = detect_f), 
  aes(x = time_taken, fill = Detected, group = Detected)) +  
  scale_fill_manual(values = detect_colours) +
  geom_histogram(binwidth = 5, boundary = 0, position = "fill") +
  facet_grid(type ~ trend) +
  labs(x = "Time taken (Seconds)", y = "Proportion of evaluations")

@@

>>>>

```{r hist_height,  eval=FALSE, fig.cap = "Six histograms show the distribution of the time taken to submit a response for each combination of trend and type of display. The time taken to evaluate each display is broken into five second windows. The height of the histogram bars show how many evaulations were submitted within each time window. The orange regions of the bars show the amount of correct detections, and the green regions are the amount of incorrect detections."}
ggplot(d %>% rename(Detected = detect_f), 
  aes(x = time_taken, fill = Detected, group = Detected)) + 
  scale_fill_manual(values = detect_colours) +
  geom_histogram(binwidth = 5, boundary = 0) +
  facet_grid(type ~ trend) + 
  theme(legend.position = "bottom") +
  labs(x = "Time taken (Seconds)", y = "Amount of evaluations")
```

```{r hist_fill,  eval=FALSE, fig.cap = "The time taken to evaluate each display is broken into five second windows. The height of the histogram bars show the proportion of evaulations that were submitted within each time window."}
ggplot(d %>% rename(Detected = detect_f), 
  aes(x = time_taken, fill = Detected, group = Detected)) +  
  scale_fill_manual(values = detect_colours) +
  geom_histogram(binwidth = 5, boundary = 0, position = "fill") +
  facet_grid(type ~ trend) +
  labs(x = "Time taken (Seconds)", y = "Proportion of evaluations")

@@

```{r}
# Di playing
s <- d %>% group_by(type, trend, detect) %>%
  summarise(m=median(time_taken), 
            q1=quantile(time_taken, 0.25), 
            q3=quantile(time_taken, 0.75))
library(ggbeeswarm)
ggplot() + 
  geom_quasirandom(data=d, aes(x=type, y=time_taken), alpha=0.9) + 
  #geom_hline(data=s, aes(yintercept=m, colour=type)) +
  geom_point(data=s, aes(x=type, y=m, colour=type), size=5, alpha=0.7) +
  #geom_errorbar(data=s, aes(x=type, ymin=q1, ymax=q3, colour=type), width=0.3, size=2) +
  scale_colour_brewer(palette = "Dark2") +
  facet_grid(trend~detect)
```

### 2020-01-07 paper/paper.tex

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The abstract goes here. On multiple lines eventually.
\end{abstract}

@@

Geospatial statistics are often presented on the geographic map base. A
choropleth map is the common display to present aggreagated statistics
for geographic units, and they are often used to present statistics
regarding the population. Creating a choropleth map involves drawing the
administrative boundaries and filling them with colour to communicate
the value of the statistic. In Australia, there are sets of
administrative boundaries that define subdivisions of the population at
various granularities. The set of Australia statistical areas presents
an example of a heterogrenous distribution of area. The rural
communicates on a much larger geographic space than small inner city
communities. This has the negative effect of incorrectly showing the
spatial distribution of the statistic, especially when a spatial
distribution is related to the size of the areas, or the population
density.

An alternative display can also be used to effectively communicate a
spatial distribution for a set of heterogeneous areas. Viewers of
spatial distributions may come to incorrect conclusions.

@@

The Australian Cancer Atlas is an online interactive webtool created to
explore the burden of cancer on Australian communities. There are many
cancer types presented, and they can be explored on an individual or
aggregate level. The Australian communities are examined at the
Statistical Areas at Level 2 (SA2) (``Australian Statistical Geography
Standard (ASGS)'' \protect\hyperlink{ref-abs2016}{2018}) used by the
Australian Bureau of Statistics. Bayesian spatial smoothing has been
applied to incorporate the value of the statistics of the neighbouring
areas. Spatial smoothing allows for the protection privacy and gives
stability to the estimates. The statistics that can be mapped are the
diagnoses (Standardised Incidence Rates) and excess deaths for each SA2,
communicated as the difference from the Australian average of the
statistics. The values of the statistic for each geographic area is
communicated through the colour used to fill the areas, the colours are
chosen for the diverging colour scheme.

The Australian Cancer Atlas communicates the trends in the distributions
of cancer over geographic space. It uses a choropleth map display and
diverging colour scheme to draw attention to relationships between
neighbouring areas.

@@

Spatial visualisations communicate the distribution of statistics over
geographic space. The most common display for spatial data is the
choropleth map. However, the issue of using a choropleth map base
becomes obvious when considering the distributions. Map creators have
the ability to present spatial statistics in alternative displays that
can highlight the population. This work aims to show that a hexagon tile
map display is a viable alternative to the geographic map base for
presenting population statistics. The same data were shown on a
choropleth map, and on a hexagon tile map. Comparing the results of
participants who see the choropleth to those who see a hexagon tile map
will show that population related distributions are spotted more
frequently in a hexagon tile map display.

When presenting population statistics on a geographic map base, the size
of the regions can allow errornous conclusions to be drawn about the
state of the statistic over the entire population. This occurs as large
regions filled with a consistent colour or pettern can draw the
attention of map readers, and small regions are not paid equal
attention. A choropleth map is not the only display that can be used for
presenting geospatial data. Alternative maps include various cartograms,
and tesselated tile maps. They allow other variables to be included in
the display to highlight the staistical values of various geographic
areas.

@@

where \(y_{ij} = 0, 1\) whether the subject detected the data plot,
\(\mu\) is the overall mean, \(\tau_i, i=1,2\) is the map type effect,
\(\delta_j\) is the trend model effect. We are allowing for an
interaction between map type and trend model. Because the response is
binary, a logistic model is used.

@@

\includegraphics{paper_files/figure-latex/detection_compare-1.pdf} The
detection rates are usually higher when participants viewed the hexagon
tile map displays. The difference was greater when viewing a three
cities trend model. When the trend model affected all cities
participants also detected the data display in the lineup of null plots
more often. The North West to South East data plot was detected with
similar rates when viewing the choropleth and hexagon tile map displays,
though for three of the replicates the participants detected the data
display more often.

\begin{tabular}{l|l|r|r|r}
\hline
trend & type & replicate & mean & std.dev\\
\hline
NW-SE & Choro. & 1 & 25.00 & 16.41\\
\hline
NW-SE & Choro. & 2 & 21.12 & 15.28\\
\hline
NW-SE & Choro. & 3 & 21.12 & 17.05\\
\hline
NW-SE & Choro. & 4 & 21.20 & 14.52\\
\hline
NW-SE & Hex. & 1 & 19.53 & 14.56\\
\hline
NW-SE & Hex. & 2 & 21.95 & 16.42\\
\hline
NW-SE & Hex. & 3 & 22.02 & 16.12\\
\hline
NW-SE & Hex. & 4 & 18.17 & 14.51\\
\hline
three cities & Choro. & 1 & 15.30 & 13.88\\
\hline
three cities & Choro. & 2 & 25.22 & 14.36\\
\hline
three cities & Choro. & 3 & 22.49 & 15.87\\
\hline
three cities & Choro. & 4 & 18.68 & 13.74\\
\hline
three cities & Hex. & 1 & 20.37 & 17.02\\
\hline
three cities & Hex. & 2 & 19.95 & 16.59\\
\hline
three cities & Hex. & 3 & 20.02 & 17.19\\
\hline
three cities & Hex. & 4 & 18.02 & 16.93\\
\hline
all cities & Choro. & 1 & 20.14 & 16.56\\
\hline
all cities & Choro. & 2 & 22.75 & 14.72\\
\hline
all cities & Choro. & 3 & 19.88 & 15.03\\
\hline
all cities & Choro. & 4 & 19.33 & 16.15\\
\hline
all cities & Hex. & 1 & 19.83 & 15.87\\
\hline
all cities & Hex. & 2 & 16.96 & 14.65\\
\hline
all cities & Hex. & 3 & 19.21 & 15.27\\
\hline
all cities & Hex. & 4 & 20.03 & 15.11\\
\hline
\end{tabular}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/hist_height-1.pdf}
\caption{The time taken to evaluate each display is broken into five
second windows. The height of the histogram bars show how many
evaulations were submitted within each time window. The distributions
for the choropleth and hexagon tile maps are very similar. Both have a
large peak at 0-5 seconds, and then a secondary peak at 10-20 seconds.
No response took over one minute.}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/hist_fill-1.pdf}
\caption{The time taken to evaluate each display is broken into five
second windows. The height of the histogram bars show the proportion of
evaulations that were submitted within each time window.}
\end{figure}

@@

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{The amount of times each level of certainty was chosen by
participants when viewing hexagon tile map or choropleth displays.
Participants were more likely to choose a high certainty when
considering a Choropleth map. The default certainty of 3 was chosen most
for the Hexagon tile map displays.}
\end{figure}

Certainty levels are measured on a five point scale---they are
subjective assessments by the participant `how certain are you about
your choice?'.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{The most common reason for choice of plot when looking at each
trend model shown in Choropleth and Hexagon Tile maps. Clusters were the
most common reason when viewing a Hexagon Tile map, trend was the most
common choice for choropleth displays except for the all cities
display.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{The probablity of detection acheived by the contributors in
each group is shown by the points. Group B has a larger range and a
smaller inter-quartile range. Group A and both had 3 people who did not
find any of the data maps in the displays.}
\end{figure}

\hypertarget{reason-1}{%
\subsection{Reason}\label{reason-1}}

\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/choices-1.pdf}
\caption{Each facet is associated with one lineup, the height of the
points show the proportion of the participants that made each choice
when considering each lineup. The points coloureddark blue show the map
which contained a trend model, these are the correct choices. The
numbers differentiate the replicates of each trend model and type of map
display. Participants were able to select 0 to indicate they did not
want to choose a map.}
\end{figure}

The choices made by participants are examined in Figure
\ref{fig:choices}. Participants were misled by the choropleth display,
but not the hexagon display for all cities displays except (2). The maps
with a North West to South East trend was chosen with much greater
frequency in all displays. All of three cities displays, except (4),
were detected in the hexagon display. All except one lineup had at least
one participant select the correct map in the lineup as shown in Figure
\ref{fig:choices}.

@@

A generalized linear mixed effects model can account for each individual
participants' abilities as it includes a subject-specific random
intercept. As each participant provides results from 12 lineups.

This includes the two main effects map type and trend model, which gives
the fixed effects part of the model to be

where \(y_{ij} = 0, 1\) whether the subject detected the data plot,
\(\mu\) is the overall mean, \(\tau_i, i=1,2\) is the map type effect,
\(\delta_j\) is the trend model effect. We are allowing for an
interaction between map type and trend model. Because the response is
binary, a logistic model is used.

@@

\begin{tabular}{l|r|r|r|r|l}
\hline
\_ & terms & Estimate & Error & p.value & sig\\
\hline
Design & Choro & 0.0217400 & 0.1474507 & 0.8828 & \\
\hline
 & Hex & 0.4200928 & 0.2110904 & 0.0466 & *\\
\hline
Trend & Three cities & -3.2519793 & 0.4124809 & 0.0000 & ***\\
\hline
 & All cities & -1.2398974 & 0.2293327 & 0.0000 & ***\\
\hline
Interactions & Hex : Three cities & 2.3683138 & 0.4645163 & 0.0000 & ***\\
\hline
 & Hex : All cities & 1.0825772 & 0.3123987 & 0.0005 & ***\\
\hline
\end{tabular}

\begin{verbatim}
##      
##        No Yes
##   No  431 121
##   Yes 242 310
\end{verbatim}

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}

@@

\begin{verbatim}
## # A tibble: 8 x 5
##   term                       estimate std.error statistic group      
##   <chr>                         <dbl>     <dbl>     <dbl> <chr>      
## 1 (Intercept)                   0.505    0.0340     14.9  fixed      
## 2 typeHex.                      0.103    0.0448      2.31 fixed      
## 3 trendthree cities            -0.467    0.0448    -10.4  fixed      
## 4 trendall cities              -0.277    0.0448     -6.19 fixed      
## 5 typeHex.:trendthree cities    0.250    0.0633      3.95 fixed      
## 6 typeHex.:trendall cities      0.239    0.0633      3.77 fixed      
## 7 sd_(Intercept).contributor    0.118   NA          NA    contributor
## 8 sd_Observation.Residual       0.430   NA          NA    Residual
\end{verbatim}

\begin{verbatim}
##      
##        No Yes
##   No   49   1
##   Yes 624 430
\end{verbatim}

@@

\begin{verbatim}
##      
##        No Yes
##   No    0   0
##   Yes 673 431
\end{verbatim}

\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\begin{verbatim}
## Linear mixed model fit by REML ['lmerMod']
## Formula: detect ~ type * trend + (1 | contributor)
##    Data: d
## REML criterion at convergence: 1350.719
## Random effects:
##  Groups      Name        Std.Dev.
##  contributor (Intercept) 0.1179  
##  Residual                0.4296  
## Number of obs: 1104, groups:  contributor, 92
## Fixed Effects:
##                (Intercept)                    typeHex.  
##                     0.5054                      0.1033  
##          trendthree cities             trendall cities  
##                    -0.4674                     -0.2772  
## typeHex.:trendthree cities    typeHex.:trendall cities  
##                     0.2500                      0.2391
\end{verbatim}

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Choosing a visualisation method may seem to be a simple task. The
choropleth map display is commonly used for communicating spatial data.
The visualisation method used to present geospatial data will influence
the information derived by map users. Choosing an alternative display
will influence the communication of the distribution. The hexagon tile
map is presented as an alternative display. Visual inference is used to
measure the power of design. The choropleth is used as a comparison
point. The hexagon tile map display is an effective visualisation for
communicatin=g population related distributions. The hexagon tile map
display also allowed identification of the geographic distribution with
values monotonically decreasing from North West toward South East areas
of Australia.
\end{abstract}

@@

A geographic map base is commonly used to present geospatial statistics.
The statistics presented may also be aggreagated statistics for
geographic units, in this case the choropleth map is the common display
used. Choropleth maps are often used to present statistics regarding the
population. Creating a choropleth map involves drawing the
administrative boundaries and filling them with colour to communicate
the value of the statistic. The several sets of Australian statistical
areas presents an example of a heterogrenous distribution. The rural
communities in Australia operate on a much larger geographic space than
small inner city communities. This has the negative effect of
incorrectly showing the spatial distribution of the statistic,
especially when a spatial distribution is related to the size of the
areas, or the population density. An alternative display can be used to
effectively communicate a spatial distribution for a set of
heterogeneous areas.

@@

The Australian Cancer Atlas is an online interactive web tool created to
explore the burden of cancer on Australian communities. There are many
cancer types presented, and they can be explored on an individual or
aggregate level. The Australian communities are examined at the
Statistical Areas at Level 2 (SA2) (``Australian Statistical Geography
Standard (ASGS)'' \protect\hyperlink{ref-abs2016}{2018}) used by the
Australian Bureau of Statistics. Bayesian spatial smoothing has been
applied to incorporate the value of the statistics of the neighbouring
areas. Spatial smoothing allows for the protection privacy and gives
stability to the estimates. The statistics mapped are the diagnoses
(Standardised Incidence Rates) and excess deaths for each SA2,
communicated as the difference from the Australian average of the
statistics. The values of the statistic for each geographic area is
communicated through the colour used to fill the areas, the colours are
chosen for the diverging colour scheme. The Australian Cancer Atlas
communicates the trends in the distributions of cancer over geographic
space. It uses a choropleth map display and diverging colour scheme to
draw attention to relationships between neighbouring areas.

@@

Spatial visualisations communicate the distribution of statistics over
geographic space. The most common display for spatial data is the
choropleth map. However, the issue of using a choropleth map base
becomes obvious when considering population related distributions. Map
creators have the ability to present spatial statistics in alternative
displays that can highlight the population. This work aims to show that
a hexagon tile map display is a viable alternative to the geographic map
base for presenting population statistics. When presenting population
statistics on a geographic map base, the size of the regions can allow
errornous conclusions to be drawn about the state of the statistic over
the entire population. This occurs as large regions filled with a
consistent colour or pettern can draw the attention of map readers, and
small regions are not paid equal attention. A choropleth map is not the
only display that can be used for presenting geospatial data.
Alternative maps include various cartograms, and tesselated tile maps.
They allow other variables to be included in the display to highlight
the staistical values of various geographic areas.

A cartogram transforms the geographic map base. The transformation
begins with a choropleth map base, and the associated populations. There
are several algorithms for transforming, they all involve shifting
boundaries to increase or descrease the area of the geographic units on
the map depending on the statistics. The shifts of the boundaries result
in cartograms with accurate map area for each of the geographic units.
The transformation will keep boundaries of neighbouring areas connected,
but the result is unfamiliar shapes for each geographic unit.

Alternative displays make the trade off between familiar shapes and
representation of variables using the area of geographic units. The
non-contiguous cartgoram method keep the shapes of geographic units
intact, and change the size to communicate the values of statistics.
This method breaks the spatial relationship between areas, and
emphasises the difference between the statistic and the amount of land
for each geographic unit. The Dorling cartogram is another method that
does not retain the boundaries of geographic units. Each unit becomes
one circle on the cartogram dispaly, these can be sized according to the
value of a variable. Their neighbours can be illustrated by connecting
the boundaries of the circles. These cartogram methods all allow two
variables to be shown in the single display.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/liver-1.pdf}
\caption{The smoothed average of liver cancer diagnoses for Australian
males. The divering colour scheme shows dark blue areas with much lower
than average diagnoses, yellow areas with diagnoses around the
Australian average, orange with above average and red shows diagnoses
much higher than average.}
\end{figure}

Australia is an extreme case of heterogenous geographic units. To
communicate information regarding the small inner city geographic units
of Australian cities an alternative display must be employed. This will
emphasise the value of the statistic that is relating to areas with high
population levels. Figure @ref(fig:liver) shows the distribution of the
smoothed average amount of Liver cancer diagnoses for Australian Males
from XXX years. The hexagon tile map (b) shows orange hexagons for the
inner city Statistical Areas at Level 2 with higher than average levels
of diagnoses in the capital cities of Melbourne and Sydney. These inner
city areas are not visible on the choropleth map (a). There are still
many light blue areas in both displays. The northern Queensland, and
Northern Territory SA2's with higher than average rates are still
visible but utilise less map space on the hexagon tile map.

@@

where \(y_{ij} = 0, 1\) whether the subject detected the data plot,
\(\mu\) is the overall mean, \(\tau_i, i=1,2\) is the map type effect,
\(\delta_j\) is the trend model effect. We are allowing for an
interaction between map type and trend model. As the response for
whether the participant detected the data plot is binary, a logistic
model is used.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/detection_compare-1.pdf}
\caption{The detection rates achieved by participants are contrasted
when viewing the four replicates of the three trend models. Each point
shows the probability of detection for the lineup display, the facets
separate the trend models hidden in the lineup. The points for the same
data set shown in a choroleth or haxgon tile map display are linked to
show the difference in the detection rate.}
\end{figure}

The detection rates are usually higher when participants viewed the
hexagon tile map displays. The difference was greater when viewing a
three cities trend model. When the trend model affected all cities
participants also detected the data display in the lineup of null plots
more often. The North West to South East data plot was detected with
similar rates when viewing the choropleth and hexagon tile map displays,
though for three of the replicates the participants detected the data
display more often.

@@

The amount of time taken for a participants to submit a response is
explored. The time considered was the amount of time it took for a
participant to submit their choice of plot, the reason for their choice,
and the perceived difficulty of the choice. The time was taken when each
participant hit ``Next''.

The relationship between time taken to submit, and the correct detection
by the participant of the real data display in a lineup is considered in
Figure @ref(fig:hist\_height). The bimodal distributions are similar
across the type of display and the trend model hidden in the real data
plot. The hexagon map displays have a peak for the group that took up to
5 seconds to submit a response. When viewing a choropleth map display,
the second peak occured at around 20 seconds. The longest a participant
took to respond was 60 seconds.

@@

The proportion of correct and incorrect responses within each 5 second
time window. For the NW-SE distribution, it is equally likely that quick
responders will be correct or incorrect. It is more likely that
participants picked the correct response if they took longer to respond.
For the Choropelth

However, these quick responders were likely to be either correct or
incorrect. The quick responders viewing a choropleth display were
unlikely to choose correctly, unless viewing a NW-SE trend model.

The choropleth map displays that included a distribution of three cities
was very difficult for participants.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/certainty-1.pdf}
\caption{The amount of times each level of certainty was chosen by
participants when viewing hexagon tile map or choropleth displays.
Participants were more likely to choose a high certainty when
considering a Choropleth map. The mid value of 3 was the default
certainty, it was chosen most for the Hexagon tile map displays.}
\end{figure}

Certainty levels are measured on a five point scale---they are
subjective assessments by the participant `How certain are you about
your choice?'.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/reason-1.pdf}
\caption{The amount of participants that selected each reason for their
choice of plot when looking at each trend model shown in Choropleth and
Hexagon Tile maps. The facets show whether or not the choice was
correct.}
\end{figure}

The most common choice made by participant was clusters, this response
was expected when participants viewed the Hexagon display of the all
cities and three cities distributions.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/contributors-1.pdf}
\caption{The probablity of detection acheived by the contributors in
each group is shown by the points. Group B has a larger range and a
smaller inter-quartile range. Group A and both had 3 people who did not
find any of the data maps in the displays.}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/choices-1.pdf}
\caption{Each facet is associated with one lineup, the height of the
points show the proportion of the participants that made each choice
when considering each lineup. The points coloured orange show the map
which contained a trend model, these are the correct choices. The
numbers differentiate the replicates of each trend model and type of map
display. Participants were able to select 0 to indicate they did not
want to choose a map.}
\end{figure}

The choices made by participants are examined in Figure
@ref(fig:choices). Participants were misled by the choropleth display,
but not the hexagon display for all cities displays except (2). The maps
with a North West to South East trend was chosen with much greater
frequency in all displays. All of three cities displays, except (4),
were detected in the hexagon display. All except one lineup had at least
one participant select the correct map in the lineup as shown in Figure
@ref(fig:choices\}.

@@

The likelihood of detecting the data plot in the lineup can be modelled
using a generalized linear mixed effects model. The base R (R Core Team
\protect\hyperlink{ref-RCore}{2019}) \texttt{glm()} function implements
generalised linear models. The model used includes the two main effects
map type and trend model, which gives the fixed effects model to be:

\[\widehat{y_{ij}} = \mu + \tau_i + \delta_j + (\tau\delta)_{ij} + \epsilon_{i,j}, ~~~ i=1,2; ~j=1,2,3\]

where \(y_{ij} = 0, 1\) whether the subject detected the data plot,
\(\mu\) is the overall mean, \(\tau_i, i=1,2\) is the map type effect,
\(\delta_j\) is the trend model effect. We are allowing for an
interaction between map type and trend model. Because the response is
binary, a logistic model is used. This model can account for each
individual participants' abilities as it includes a subject-specific
random intercept. As each participant provides results from 12 lineups.

@@

The model specifies a logistic link, this means the predicted values
from the \texttt{glm} model should be back-transformed to fit between 0
and 1. They are transformed with the link specified below:

\[\mu = \frac{e^{\eta}}{1 + e^{\eta}}\] \[\eta = f(\tau_i,\delta_j)\]

\begin{tabular}{l|r|r|r|r|l|l}
\hline
\_ & terms & Estimate & Error & Stat & p.value & sig\\
\hline
Design & Intercept & 0.0217400 & 0.1474507 & 0.1474391 & 0.8828 & \\
\hline
 & Hex & 0.4200928 & 0.2110904 & 1.9901083 & 0.0466 & *\\
\hline
Trend & Three cities & -3.2519793 & 0.4124809 & -7.8839514 & 0.0000 & ***\\
\hline
 & All cities & -1.2398974 & 0.2293327 & -5.4065450 & 0.0000 & ***\\
\hline
Interactions & Hex : Three cities & 2.3683138 & 0.4645163 & 5.0984512 & 0.0000 & ***\\
\hline
 & Hex : All cities & 1.0825772 & 0.3123987 & 3.4653698 & 0.0005 & ***\\
\hline
\end{tabular}

\begin{verbatim}
##                 
##                  Detected? No Detected? Yes
##   Predicted: No           431           121
##   Predicted: Yes          242           310
\end{verbatim}

@@

The estimated marginal mean is the fitted value from a model over the
treatment effects. The treatment effects considered in this model are
map type displayed, trend model included, and the replicate. The
\texttt{emmeans} package is used to calculate the estimated marginal
mean. The value, between 0 and 1, can be interpretted as the probability
of detecting the real data plot in the lineup, for each combination of
map type, trend model and replicate.

\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}

@@

\begin{verbatim}
## # A tibble: 8 x 5
##   term                       estimate std.error statistic group      
##   <chr>                         <dbl>     <dbl>     <dbl> <chr>      
## 1 (Intercept)                   0.505    0.0340     14.9  fixed      
## 2 typeHex.                      0.103    0.0448      2.31 fixed      
## 3 trendThree Cities            -0.467    0.0448    -10.4  fixed      
## 4 trendAll Cities              -0.277    0.0448     -6.19 fixed      
## 5 typeHex.:trendThree Cities    0.250    0.0633      3.95 fixed      
## 6 typeHex.:trendAll Cities      0.239    0.0633      3.77 fixed      
## 7 sd_(Intercept).contributor    0.118   NA          NA    contributor
## 8 sd_Observation.Residual       0.430   NA          NA    Residual
\end{verbatim}

\begin{verbatim}
##      
##       Detected? No Detected? Yes
##   No            49             1
##   Yes          624           430
\end{verbatim}

@@

\begin{verbatim}
##      
##       Detected? No Detected? Yes
##   No             0             0
##   Yes          673           431
\end{verbatim}

@@

\begin{verbatim}
## Linear mixed model fit by REML ['lmerMod']
## Formula: detect ~ type * trend + (1 | contributor)
##    Data: d
## REML criterion at convergence: 1350.719
## Random effects:
##  Groups      Name        Std.Dev.
##  contributor (Intercept) 0.1179  
##  Residual                0.4296  
## Number of obs: 1104, groups:  contributor, 92
## Fixed Effects:
##                (Intercept)                    typeHex.  
##                     0.5054                      0.1033  
##          trendThree Cities             trendAll Cities  
##                    -0.4674                     -0.2772  
## typeHex.:trendThree Cities    typeHex.:trendAll Cities  
##                     0.2500                      0.2391
\end{verbatim}
